---
title: "autos Backup"
author: "Zach"
date: 2018-11-23
output:
  github_document:
  toc: true
---

The following is a pre-baked analysis of the UCI `autos` dataset. Our goal is to
understand the relationship between vehicle characteristics and their 'risk', as
defined by actuarial risk adjustment (`symboling`). Actuaries assign these
values for insurance purposes, and it affects the premium a car owner pays for
insuring the vehicle (riskier vehicles tend to be more pricey to insure).

```{r setup, echo=FALSE, warning=FALSE}
library(tidyverse)
library(GGally)

## Data filename
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

## Helper functions
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
```

## Wrangling
<!-- ------------------------------------------------- -->

Hadley Wickham says that "importing data is 80% boredom and 20% screaming" --
I've done the screaming here so you don't have to!

The data are essentially in CSV format, though we'll need to do some wrangling.

```{r load_n_wrangle}
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```

The string-coded numbers are obnoxious; let's map those to integers.

```{r}
df_raw %>% pull(num_of_doors) %>% unique()
df_raw %>% pull(num_of_cylinders) %>% unique()
```

Manual fixes for some of the variables.

```{r}
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  )
```

Finally, we can analyze the data.

## Analysis
<!-- ------------------------------------------------- -->

Our objective with these data is to determine how various attributes of
automobiles affect actuarial risk. To that end, we'll follow a hierarchy of
exploratory tasks, aimed at 1. understanding the individual variables,
2. understanding relationships between the various inputs, and 3. understanding
how the inputs (predictors) affect the output (response).

### 1. One-dimensional analysis
<!-- -------------------------------------------------- -->

When coming to a new dataset, we should start with a `glimpse()` and a
`summary()`.

```{r}
df_autos %>% glimpse()
df_autos %>% summary()
```

The numeric input variables (predictors) seem to break down into

| Geometry   | `wheel_base`, `length`, `width`, `height`, `curb_weight` |
| Engine     | `engine_size`, `num_of_cylinders`, `engine_size`, `bore`, `stroke`, `compression_ratio`, `horsepower`, `peak_rpm` |
| Efficiency | `city_mpg`, `highway_mpg` |
| Monetary   | `price` |
| Categorical   | `make`, `fuel_type`, `aspiration`, `body_style`, `drive_wheels`, `engine_location` |

#### Categorical variables
<!-- ------------------------- -->

Categorical variables take specific, fixed values. For instance, `body_style`
takes the values `r df_autos %>% pull(body_style) %>% unique`. To get a sense of
each variable, we'll just report the total number of each instance in the data
set in a bar graph.

```{r eda1d_helper}
make_1d_bar <- function(varname) {
  var = enquo(varname)

  df_autos %>%
    count(!! var) %>%
    ggplot(aes(fct_reorder(!! var, n), n)) +
    geom_bar(stat = "identity") +
    theme(axis.text.x = element_text(angle = 315))
}

make_1d_hist <- function(varname, bins = 30) {
  var = enquo(varname)

  df_autos %>%
    ggplot(aes(!! var)) +
    geom_histogram(bins = bins)
}
```

```{r eda1d_make}
make_1d_bar(make)
```

We can see that we have many more models from Japanese manufacturers; attributes
of these vehicles will be over-represented in our data.

```{r eda1d_body_style}
make_1d_bar(body_style)
```

We also see far more sedans and hatchbacks than other body styles.

```{r eda1d_engine_location}
make_1d_bar(engine_location)
```

We have almost exclusively front-engine vehicles.

```{r eda1d_fuel_type}
make_1d_bar(fuel_type)
```

There are not many diesel engines in this dataset.

#### Continuous variables
<!-- ------------------------- -->

Continuous variables can take an infinite number of values; for instance 1 / n
for n = 1,2,.... Because of this, we can't do a bar graph as above. Instead, we
define a number of continuous bins into which we sort the continuous values.
This is called a 'histogram'.

```{r eda1d_mpg}
make_1d_hist(horsepower, bins = 20)
```

Rule 1 of Histograms is: Play with the bin size!

```{r eda1d_mpg}
make_1d_hist(horsepower, bins = 40)
```

There are many 'fingers' in the horsepower distribution; this is revealed by
varying the number of bins.

```{r eda1d_engine_size}
make_1d_hist(engine_size, bins = 35)
```

```{r eda1d_curb_weight}
make_1d_hist(curb_weight, bins = 30)
```

```{r eda1d_compression_ratio}
make_1d_hist(compression_ratio, bins = 30)
```

There is a fair bit of bi-modality in compression ratio! Let's follow this up
below.

```{r eda1d_peak_rpm}
make_1d_hist(peak_rpm, bins = 20)
```

```{r eda1d_city_mpg}
make_1d_hist(city_mpg, bins = 30)
```

```{r eda1d_highway_mpg}
make_1d_hist(highway_mpg, bins = 30)
```

```{r eda1d_price}
make_1d_hist(price, bins = 25)
```

Price is bounded below by zero, so it will tend to be heavily 'skewed' or
tilted. We say this distribution has a 'long right tail', as there are few
instances with very large values.

### 2. Pairwise analysis
<!-- -------------------------------------------------- -->

```{r pairs}
df_autos %>%
  ## select_if(is.numeric) %>%
  select(
    body_style,
    curb_weight,
    engine_size,
    compression_ratio,
    horsepower,
    city_mpg,
    highway_mpg,
    price
  ) %>%
  ggpairs()
```

```{r diesel_w_cr}
df_autos %>%
  ggplot(aes(compression_ratio, color = fuel_type, group = fuel_type)) +
  geom_freqpoly()
```

Diesel engines use compression ignition (rather than spark ignition), and so
tend to be of higher CR than petrol engines.

```{r horsepower_v_other}
df_autos %>%
  ggplot(aes(horsepower, engine_size)) +
  ## geom_point(aes(color = make), size = 0.5, alpha = 1 / 3) +
  geom_hex(bins = 35) +
  theme(legend.position = "bottom")
```

### 3. Predictor-response analysis
<!-- -------------------------------------------------- -->

### Fun fact
<!-- -------------------------------------------------- -->
There's one last thing to note in these data:

```{r}
df_autos %>%
  ggplot(aes(price)) +
  geom_smooth(aes(y = highway_mpg, color = "Highway")) +
  geom_smooth(aes(y = city_mpg, color = "City")) +
  geom_point(aes(y = highway_mpg, color = "Highway")) +
  geom_point(aes(y = city_mpg, color = "City")) +
  scale_x_log10() +
  scale_y_log10() +
  labs(
    x = "Price",
    y = "Efficiency"
  )
```

More expensive cars are less efficient! You get what you pay for, I suppose....
