---
title: "Intro to Visualization"
author: "Zach"
date: 2018-11-24
output:
  github_document:
    toc: true
    toc_depth: 6
---

<!-- solution-begin -->
The following is a document I use to teach visualization basics.
<!-- solution-end -->

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r parameters, message = FALSE, warning = FALSE}
# Libraries
## --------------------------------------------------
library(tidyverse)

# Parameters
## --------------------------------------------------
## Data filename
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

# Parameters
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
```

## Wrangling
<!-- ------------------------------------------------- -->

First, we need to load and wrangle the data a bit.

```{r load_n_wrangle}
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```

The string-coded numbers are obnoxious; let's map those to integers.

```{r}
df_raw %>% pull(num_of_doors) %>% unique()
df_raw %>% pull(num_of_cylinders) %>% unique()
```

Manual fixes for some of the variables.

```{r}
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  )
```

Finally, we can analyze the data.

## Visualization
<!-- ------------------------------------------------- -->

Let's work our way through some visualization concepts.

### Bar Graph
<!-- ------------------------- -->

```{r bar_graph}
df_autos %>%
  ggplot(aes(body_style)) +
  geom_bar()
```

__q1.1__ What does this plot show?

<!-- solution-begin -->
This shows the total number of cars in the data, according to their `body_style`.
<!-- solution-end -->

__q1.2__ Why are there five bars?

<!-- solution-begin -->
There are five kinds of `body_style` that appear in the data; they are `r df_autos %>% pull(body_style) %>% unique`.
<!-- solution-end -->

__q1.3__ Could we plot `length` this way?

<!-- solution-begin -->
No; there are too many different values of `length` in the data. We would need to use a different strategy
to plot `length` in this fashion.
<!-- solution-end -->

### Histograms
<!-- ------------------------- -->

```{r hist}
df_autos %>% 
  ggplot(aes(length)) +
  geom_histogram(bins = 25)
```

__q1.1__ What does this plot show?

<!-- solution-begin -->
This shows the number of cars in the data landing in particular 'bins' of `length`. Above,
we have chosen to chop the range of values of `length` up into `25` equal-spaced bins,
and plot the number of cars in each bin much like the bar graph above.
<!-- solution-end -->

__q2.2__ How does this plot differ from the bar graph above?

<!-- solution-begin -->
The bar graph plotted a discrete quantity, while the histogram plots a continuous variable.
The bar graph directly reports counts for each of its possible values, while the histogram
first creates an approximation of the continuous data.
<!-- solution-end -->

__q2.3__ What choice do we need to make when creating a histogram?

<!-- solution-begin -->
We must choose the bins. We can do this by making them equally spaced and choosing a length or
number of bins. We could also choose unequally-spaced bins.

The Cardinal Rule of Histograms: Play with the bin spacing. Make sure to try several values
of `bins` when first exploring a dataset; certain patterns only appear at particular bin
spacings. 
<!-- solution-end -->

### Pair plots
<!-- ------------------------- -->

__q3.1__ Make a prediction: How does `length` vary with `body_style`?

```{r length_v_body_style}
df_autos %>% 
  ggplot(aes(body_style, length)) +
  geom_point()
```

__q3.2__ What `body_style` tends to have the shortest car?

<!-- solution-begin -->
Hatchbacks have the smallest vehicles in the data
<!-- solution-end -->

```{r length_v_body_style}
df_autos %>% 
  ggplot(aes(body_style, length)) +
  geom_boxplot()
```

__q3.3__ What does this plot show?

<!-- solution-begin -->
This is a *boxplot*, a visual summary invented by American statistician John Tukey. It gives a measure
of the 'center' of the data (the median) as a horizontal line, two measures of the 'spread' of the
data as the white boxes and the vertical 'whiskers', and shows outliers as solid dots.

This visualization shows that wagons tend to be the longest (in terms of median), but sedans do have
longer models.
<!-- solution-end -->

### Beyond pairs
<!-- ------------------------- -->

```{r three}
df_autos %>% 
  ggplot(aes(body_style, length, color = as.factor(num_of_doors))) +
  geom_point() +
  theme(legend.position = "bottom")
```

__q4.1__ How many variables can we see in the plot above?

<!-- solution-begin -->
Three
<!-- solution-end -->

__q4.2__ How does `num_of_doors` vary with `body_style`?

<!-- solution-begin -->
All the convertibles and hardtops are two-door; hatchbacks and sedans are a (skewed) mix,
and all our wagons are four-door.
<!-- solution-end -->

__q4.3__ What does `NA` mean?

<!-- solution-begin -->
`NA` means 'not available'; this piece of data is missing! This illustrates an important point:
Data can have errors!
<!-- solution-end -->

```{r three2}
df_autos %>% 
  ggplot(aes(fuel_type, body_style)) +
  geom_point(aes(color = compression_ratio))
```

__q4.4__ Is this a good visualization?

<!-- solution-begin -->
This is a terrible visualization! Here are a few reasons why:

1. Overplotting: There are multiple points overlapping on the same position. It's impossible
   to tell this from the visualization above. A solution is to 'jitter' the data.
2. Use of color: Color is one of the least-effective ways to convey information. We could
   plot `compression_ratio` against one of the axes instead, and it would be far more
   effective.
3. Orientation: Generally, we have more space along a horizontal axis than a vertical
   one. We could put `body_style` along the horizontal axis for a more effective use
   of space. (This is a relatively minor point.)
<!-- solution-end -->

__q4.4__ Improve this visualization of `compression_ratio`, `body_style`, and `fuel_type`.

<!-- solution-begin -->
Here's my improved graph. From this, we can see the striking difference between diesel and gas engines;
diesel engines use a 'compression ignition' approach to combustion, while gas engines use spark plugs.
By design, diesel engines tend to have a higher compression ratio, in order to ignite the fuel.

```{r three3}
df_autos %>% 
  ggplot(aes(body_style, compression_ratio, color = fuel_type)) +
  geom_point() +
  theme(legend.position = "bottom")
```
<!-- solution-end -->

<!-- solution-begin -->
### Visual hierarchy
<!-- ------------------------- -->

Here's some bonus material:

There is a hierarchy of perceptual tasks, ranked in terms of our ability to
distinguish features and make comparisons. To maximize readability of a graph,
one should use the highest-ranked tasks to display the most relevant features of
the data.

1. Position along a common scale
2. Position on identical but nonaligned scale
3. Length
4. Angle or slope
5. Area
6. Volume or Density or Color saturation
7. Color hue
<!-- solution-end -->
