---
title: "Intro to Visualization"
author: "Zach"
date: 2018-11-24
output:
  github_document:
    toc: true
    toc_depth: 6
---

<!-- solution-begin -->
The following is a document I use to teach visualization basics.
<!-- solution-end -->

## Setup
(Keep this folded!)

```{r include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r parameters, message = FALSE, warning = FALSE}
# Libraries
## --------------------------------------------------
library(tidyverse)

# Parameters
## --------------------------------------------------
## Data filename
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

# Parameters
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
## Common theme settings
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    aspect.ratio = 3 / 4,

    plot.margin   = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    panel.spacing = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),

    legend.text = element_text(size = 10),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90")
  )
}
```


First, we need to load and wrangle the data a bit.

```{r load_n_wrangle, warning=FALSE, message=FALSE}
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```

The string-coded numbers are obnoxious; let's map those to integers.

```{r}
df_raw %>% pull(num_of_doors) %>% unique()
df_raw %>% pull(num_of_cylinders) %>% unique()
```

Manual fixes for some of the variables.

```{r}
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  )
```

Finally, we can analyze the data.

## Visualization
<!-- ------------------------------------------------- -->

I've got some data on automobiles. Here are the car facts we've got:

```{r}
df_autos %>% names
```

Here's a quick look at some of the data.

```{r}
df_autos %>% head
```

It's hard to tell what's going on from this! We're going to *visualize* the data to help us understand what's going on.

### Single plots
<!-- ------------------------- -->

```{r}
df_autos %>% 
  ggplot(aes(body_style, id)) +
  geom_point() +
  theme_common() +
  labs(
    y = "Car identifier index"
  )
```

__q1.1__ What do the 'dots' here represent?

<!-- solution-begin -->
Each 'dot' represents an individual car. They are organized by `body_style`, and their `id` in the dataset.
<!-- solution-end -->

__q1.2__ Which `body_style` do we have the most data on?

<!-- solution-begin -->
Maybe hatchbacks or sedans? It's a little difficult to tell from the figure above.
<!-- solution-end -->

```{r}
df_autos %>% 
  ggplot(aes(body_style)) +
  geom_bar() +
  theme_common()
```

__q1.3__ What does this plot show?

<!-- solution-begin -->
This shows the total number of cars within each `body_style`.
<!-- solution-end -->

__q1.4__ Is this easier or harder to read than the previous plot?

<!-- solution-begin -->
I *claim* that this is easier to read.
<!-- solution-end -->

### Pair plots
<!-- ------------------------- -->

```{r length_v_body_style}
df_autos %>% 
  ggplot(aes(body_style, length)) +
  geom_point() +
  theme_common()
```

__q2.1__ What does this plot show us?

<!-- solution-begin -->
The `length` of individual cars, organized by their `body_style`.
<!-- solution-end -->

__q2.2__ What `body_style` tends to have the shortest car?

<!-- solution-begin -->
Hatchbacks have the smallest vehicles in the data
<!-- solution-end -->

__q2.3__ Can you tell which `body_style` has the largest car, on average?

<!-- solution-begin -->
This is a little difficult to do from the plot above.
<!-- solution-end -->

```{r length_v_body_style}
df_autos %>% 
  group_by(body_style) %>% 
  mutate(mean_len = mean(length, na.rm = F)) %>% 
  ungroup() %>% 
  ggplot(aes(body_style, length)) +
  geom_point(aes(color = "Individual")) +
  geom_point(
    data = . %>% select(body_style, mean_len) %>% unique,
    mapping = aes(y = mean_len, color = "Average"), 
    size = 4, 
    alpha = 1 / 2
  ) +
  scale_color_discrete(name = NULL) +
  theme_common()
```

<!-- solution-begin -->
Punchline: Sometimes, summarizing the data can help us see underlying patterns.
<!-- solution-end -->

### Beyond pairs
<!-- ------------------------- -->

```{r three}
df_autos %>% 
  ggplot(aes(body_style, length, color = as.factor(num_of_doors))) +
  geom_point() +
  theme_common() +
  theme(legend.position = "bottom")
```

__q3.1__ How many variables can we see in the plot above?

<!-- solution-begin -->
Three; they are `length`, `body_style`, and `num_of_doors`.
<!-- solution-end -->

__q3.2__ How does `num_of_doors` vary with `body_style`?

<!-- solution-begin -->
All the convertibles and hardtops are two-door; hatchbacks and sedans are a (skewed) mix,
and all our wagons are four-door.
<!-- solution-end -->

__q3.3__ What does `NA` mean?

<!-- solution-begin -->
`NA` means 'not available'; this piece of data is missing! This illustrates an important point:
Data can have errors!
<!-- solution-end -->

```{r three2}
df_autos %>% 
  ggplot(aes(fuel_type, body_style)) +
  geom_point(aes(color = compression_ratio)) +
  theme_common()
```

__q3.4__ Is this a good visualization? Why or why not?

<!-- solution-begin -->
This is a terrible visualization! Here are a few reasons why:

1. Overplotting: There are multiple points overlapping on the same position. It's impossible
   to tell this from the visualization above. A solution is to 'jitter' the data.
2. Use of color: Color is one of the least-effective ways to convey information. We could
   plot `compression_ratio` against one of the axes instead, and it would be far more
   effective.
3. Orientation: Generally, we have more space along a horizontal axis than a vertical
   one. We could put `body_style` along the horizontal axis for a more effective use
   of space. (This is a relatively minor point.)
<!-- solution-end -->

__q3.4__ Improve this visualization of `compression_ratio`, `body_style`, and `fuel_type`.

<!-- solution-begin -->
Here's my improved graph. From this, we can see the striking difference between diesel and gas engines;
diesel engines use a 'compression ignition' approach to combustion, while gas engines use spark plugs.
By design, diesel engines tend to have a higher compression ratio, in order to ignite the fuel.

```{r three3}
df_autos %>% 
  ggplot(aes(body_style, compression_ratio, color = fuel_type)) +
  geom_point() +
  theme_common() +
  theme(legend.position = "bottom")
```
<!-- solution-end -->

<!-- solution-begin -->
### Visual hierarchy
<!-- ------------------------- -->

Here's some bonus material:

There is a hierarchy of perceptual tasks, ranked in terms of our ability to
distinguish features and make comparisons. To maximize readability of a graph,
one should use the highest-ranked tasks to display the most relevant features of
the data.

1. Position along a common scale
2. Position on identical but nonaligned scale
3. Length
4. Angle or slope
5. Area
6. Volume or Density or Color saturation
7. Color hue
<!-- solution-end -->
