---
title: "Judging Visualizations"
author: "Zach"
date: 2019-03-03
output:
  github_document:
    toc: true
    toc_depth: 6
---

This document uses the "Automobile Data Set" from the UCI Machine Learning
database. You will need to
[download](https://archive.ics.uci.edu/ml/datasets/Automobile) the relevant data
file and change `file_name` below to point to the appropriate directory.

<!-- solution-begin -->
The document below produces visualizations for exploring the data, and provides
discussion prompts for students to consider.

Running `make` in this directory will create a `challenge.Rmd` document without
the solution commentary. This can be useful if you want to assign this document
for students to work through on their own.
<!-- solution-end -->

```{r parameters, message = FALSE, warning = FALSE}
# Libraries
## --------------------------------------------------
library(tidyverse)

# Parameters
## --------------------------------------------------
## Data filename -- you will need to change this!
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

# Parameters
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
## Common theme settings
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 14),
    axis.text.y = element_text(size = 14),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    aspect.ratio = 3 / 4,

    plot.margin   = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    panel.spacing = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),

    legend.title = element_text(size = 14),
    legend.text = element_text(size = 14),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90")
  )
}
```

First, we need to load and wrangle the data a bit.

```{r load_n_wrangle, warning=FALSE, message=FALSE}
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```

The string-coded numbers are obnoxious; let's map those to integers.

```{r}
df_raw %>% pull(num_of_doors) %>% unique()
df_raw %>% pull(num_of_cylinders) %>% unique()
```

Manual fixes for some of the variables.

```{r}
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  ) %>%
  rowid_to_column(var = "id")
```

Finally, we can analyze the data.

## Judging
<!-- -------------------------------------------------- -->

Look at the following visualization:

```{r three_orig}
df_autos %>%
  ggplot(aes(fuel_type, body_style)) +
  geom_point(aes(color = compression_ratio), size = 8) +
  theme_common() +
  theme(legend.position = "bottom")
```

__q1__ Which rung on the visual hierarchy is used to visualize `compression_ratio` here?

<!-- solution-begin -->
`compression_ratio` is displayed here using color hue, which is on the bottom of
the visual hierarchy.
<!-- solution-end -->

__q2__ Is this a good visualization? Why or why not?

<!-- solution-begin -->
This is a terrible visualization! Here are a few reasons why:

1. Overplotting: There are multiple points overlapping on the same position. It's impossible
   to tell this from the visualization above. A solution is to 'jitter' the data.
2. Use of color: Color is one of the least-effective ways to convey information. We could
   plot `compression_ratio` against one of the axes instead, and it would be far more
   effective.
3. Orientation: Generally, we have more space along a horizontal axis than a vertical
   one. We could put `body_style` along the horizontal axis for a more effective use
   of space. (This is a relatively minor point.)
<!-- solution-end -->

__q3__ Construct a *quality* visualization of `compression_ratio`, `body_style`, and `fuel_type`.

Hint: Sometimes students can have a hard time making concrete suggestions. In
this case, ask students to *swap* pairs of variables. Have them comment on what
is better (or worse) about intermediate visualizations.

```{r three_repeat}
df_autos %>%
  ggplot(aes(fuel_type, body_style)) +
  geom_point(aes(color = compression_ratio), size = 8) +
  theme_common() +
  theme(legend.position = "bottom")
```

<!-- solution-begin -->
Here's my improved graph. From this, we can see the striking difference between
diesel and gas engines; diesel engines use a 'compression ignition' approach to
combustion, while gas engines use spark plugs. By design, diesel engines tend to
have a higher compression ratio, in order to ignite the fuel.

```{r three_improved}
df_autos %>%
  ggplot(aes(body_style, compression_ratio, color = fuel_type)) +
  geom_point(size = 3) +
  theme_common() +
  theme(legend.position = "bottom")
```

From this visualization, it's immediately apparent that gas and diesel engines
have very different `compression_ratio` values.
<!-- solution-end -->

__q4__ What can you see in your improved visualization?

<!-- solution-begin -->
Some observations, in no particular order:

1. Based on the density of dots, there are more `hatchback` and `sedan` style
   vehicles in this dataset, compared to the other types.
2. `diesel` engines tend to have a much higher compression ratio than `gas`
   engines. This makes sense -- diesel engines use compression ignition to
   ignite the fuel, rather than a sparkplug. We might expect a higher
   compression ratio as a consequence.
3. There are no `diesel` convertibles. From some quick Google searching, it
   seems that diesel cars used to be noisy and smelly, which is in conflict with
   the perceived luxury of convertibles. Modern diesel convertibles apparently
   exist.
4. There are far fewer `diesel` vehicles than `gas` vehicles in this dataset. A
   Finnish friend of mine pointed out that diesel vehicles are unpopular in the
   US, but are more popular in Europe.
<!-- solution-end -->

## More Graphs
<!-- -------------------------------------------------- -->

```{r mpg_base}
df_autos %>%
  ggplot(aes(city_mpg, highway_mpg)) +
  geom_point(point = 3) +
  theme_common()
```

__q5__ What does the following visualization show us?

<!-- solution-begin -->
From this, we can see:

- Highway and city mileage are closely related.
- Highway mileage is generally larger than city mileage.
<!-- solution-end -->

__q6__ Is this a good visualization? Why or why not?

<!-- solution-begin -->
This visualization is ok, but due to the way the data are plotted,
we're missing some facts:

- This doesn't show us *all* the data; there is overplotting. This is because
  MPG is generally rounded.
- Some of the cars have exactly the same highway and city mileage; this is
  difficult to tell from the current vis.
<!-- solution-end -->

__q7__ How might we improve this graph?

<!-- solution-begin -->
Here's my modified graph:

```{r mpg_improved}
df_autos %>%
  ggplot(aes(city_mpg, highway_mpg)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_count() +
  scale_size_continuous(name = "Count") +
  theme_common() +
  theme(legend.position = "bottom")
```
<!-- solution-end -->

__q8__ What can you see in your improved visualization?

<!-- solution-begin -->
What this shows us:

1. *All* of these cars have greater (or equal) highway than city MPG.
2. An explicit guide for the line of equal city and highway mileage. This
   shows us that a few cars have the same mileage for both kinds of driving.
3. The number of vehicles at each point -- MPG is concentrated towards the
   middle-lower range.
<!-- solution-end -->
