---
title: "Judging Visualizations"
author: "Zach"
date: 2019-03-03
output:
  github_document:
    toc: true
    toc_depth: 6
---

```{r parameters, message = FALSE, warning = FALSE}
# Libraries
## --------------------------------------------------
library(tidyverse)

# Parameters
## --------------------------------------------------
## Data filename
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

# Parameters
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
## Common theme settings
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    aspect.ratio = 3 / 4,

    plot.margin   = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    panel.spacing = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),

    legend.text = element_text(size = 10),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90")
  )
}
```

First, we need to load and wrangle the data a bit.

```{r load_n_wrangle, warning=FALSE, message=FALSE}
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```

The string-coded numbers are obnoxious; let's map those to integers.

```{r}
df_raw %>% pull(num_of_doors) %>% unique()
df_raw %>% pull(num_of_cylinders) %>% unique()
```

Manual fixes for some of the variables.

```{r}
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  ) %>%
  rowid_to_column(var = "id")
```

Finally, we can analyze the data.

## Brief Overview of the Data
<!-- -------------------------------------------------- -->

Let's get an overview of the data. It's always a good idea to look at a glimpse
and a summary.

```{r data_glimpse}
df_autos %>% glimpse
```

```{r data_summary}
df_autos %>% summary
```

As a reminder, we usually start with single-variable investigations.

```{r hist_price}
df_autos %>%
  ggplot(aes(price)) +
  geom_histogram(bins = 30) +
  theme_common() +
  labs(
    x = "Price (1985 Dollars)",
    y = "Count"
  )
```

We then move to 2D investigations.

```{r price_v_weight}
df_autos %>%
  ggplot(aes(curb_weight, price)) +
  geom_point() +
  scale_x_log10() +
  scale_y_log10() +
  theme_common() +
  labs(
    x = "Curb Weight (lbs)",
    y =  "Price (1985 Dollars)"
  )
```

## Judging
<!-- -------------------------------------------------- -->

Look at the following visualization:

```{r three2}
df_autos %>%
  ggplot(aes(fuel_type, body_style)) +
  geom_point(aes(color = compression_ratio)) +
  theme_common()
```

__q1__ Which rung on the visual hierarchy is used to visualize `compression_ratio` here?

<!-- solution-begin -->
`compression_ratio` is displayed here using color hue, which is on the bottom of
the visual hierarchy.
<!-- solution-end -->

__q2__ Is this a good visualization? Why or why not?

<!-- solution-begin -->
This is a terrible visualization! Here are a few reasons why:

1. Overplotting: There are multiple points overlapping on the same position. It's impossible
   to tell this from the visualization above. A solution is to 'jitter' the data.
2. Use of color: Color is one of the least-effective ways to convey information. We could
   plot `compression_ratio` against one of the axes instead, and it would be far more
   effective.
3. Orientation: Generally, we have more space along a horizontal axis than a vertical
   one. We could put `body_style` along the horizontal axis for a more effective use
   of space. (This is a relatively minor point.)
<!-- solution-end -->

__q3__ Construct a *quality* visualization of `compression_ratio`, `body_style`, and `fuel_type`.

<!-- solution-begin -->
Here's my improved graph. From this, we can see the striking difference between
diesel and gas engines; diesel engines use a 'compression ignition' approach to
combustion, while gas engines use spark plugs. By design, diesel engines tend to
have a higher compression ratio, in order to ignite the fuel.

```{r three3}
df_autos %>%
  ggplot(aes(body_style, compression_ratio, color = fuel_type)) +
  geom_point() +
  theme_common() +
  theme(legend.position = "bottom")
```

From this visualization, it's immediately apparent that gas and diesel engines
have very different `compression_ratio` values.
<!-- solution-end -->

## More Graphs
<!-- -------------------------------------------------- -->

```{r mpg_base}
df_autos %>%
  ggplot(aes(city_mpg, highway_mpg)) +
  geom_point() +
  theme_common()
```

__q4__ What does the following visualization show us?

<!-- solution-begin -->
From this, we can see:

- Highway and city mileage are closely related.
- Highway mileage is generally larger than city mileage.

We're missing some things, though!

- This doesn't show us *all* the data; there is overplotting. This is because
  MPG is generally rounded.
<!-- solution-end -->

__q5__ How might we improve this graph?

<!-- solution-begin -->
Here's my modified graph:

```{r mpg_improved}
df_autos %>%
  ggplot(aes(city_mpg, highway_mpg)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2) +
  geom_count() +
  scale_size_continuous(name = "Count") +
  theme_common() +
  theme(legend.position = "bottom")
```

What this shows us:

- *All* of these cars have greater (or equal) highway than city MPG.
- An explicit guide for the line of equal city and highway mileage. This
  shows us that a few cars have exactly the same mileage!
- The number of vehicles at each point -- MPG is concentrated towards the
  middle-lower range.
<!-- solution-end -->
