---
title: "The Visual Hierarchy"
author: "Your Name"
date: 
bibliography: bibtex_database.bib
output: github_document
---

In this lesson, we will illustrate different methods of visualizing information.
The purpose of this exercise is to determine which methods are the most
effective -- in terms of 'ease of interpretation.' You will rank a number of
different visualizations -- of the same data -- in terms of how easily you can
draw comparisons. This will lead to prescriptive suggestions about constructing
visualizations, which will be useful when designing graphics.


```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)

sel_wheels <- c("fwd", "rwd")
sel_style  <- c("hatchback", "sedan", "wagon")

asp_ratio  <- 4 / 4
img_width  <- 5
img_height <- img_width * asp_ratio

# Common theme settings
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    aspect.ratio = asp_ratio,

    ## plot.margin   = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    ## panel.spacing = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),

    legend.text = element_text(size = 10),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90")
  )
}

# Custom save
mysave <- function(filename) {
  ggsave(
    filename = filename,
    height   = img_height,
    width    = img_width,
    units    = "in"
    ## device   = cairo_ps
  )
}
```

```{r load_n_wrangle, echo=FALSE, include=FALSE, warning=FALSE}
## Data load + wrangle.
# Parameters
## --------------------------------------------------
## Data filename
file_name <- "~/Sync/data/autos/imports-85.data"
## Variable names from the UCI site
var_names <-
  c(
    "symboling",          # Actuarial risk adjustment: -3 Safest, +3 Riskiest
    "normalized_losses",  # Average insurance loss, averaged over body_style
    "make",               # Manufacturer
    "fuel_type",          # Gas or diesel
    "aspiration",         # Standard or turbo
    "num_of_doors",       #
    "body_style",         # hardtop, wagon, sedan, hatchback, convertible
    "drive_wheels",       # Front, rear, all
    "engine_location",    # Front, rear
    "wheel_base",         # Distance between front & rear wheels
    "length",             #
    "width",              #
    "height",             #
    "curb_weight",        # Weight w/o cargo
    "engine_type",        # Engine layout
    "num_of_cylinders",   #
    "engine_size",        # Total engine displacement
    "fuel_system",        # Fuel feed system
    "bore",               # Cylinder diameter
    "stroke",             # Cylinder displacement distance
    "compression_ratio",  # Swept volume compression ratio
    "horsepower",         # Engine power
    "peak_rpm",           # Peak rotations per minute (redline?)
    "city_mpg",           # City fuel efficiency
    "highway_mpg",        # Highway fuel efficiency
    "price"               # MSRP
  )

# Parameters
## --------------------------------------------------
parse_numchar <- function(x) {
  case_when(
        x == "two" ~ 2L,
        x == "three" ~ 3L,
        x == "four" ~ 4L,
        x == "five" ~ 5L,
        x == "six" ~ 6L,
        x == "seven" ~ 7L,
        x == "eight" ~ 8L,
        x == "nine" ~ 9L,
        x == "ten" ~ 10L,
        x == "eleven" ~ 11L,
        x == "twelve" ~ 12L,
        TRUE ~ NA_integer_
      )
}
```

```{r rename, echo=FALSE, warning=FALSE, message=FALSE}
## First, we need to load and wrangle the data a bit.
df_raw <-
  read_csv(
    file_name,
    col_names = var_names,
    col_types = cols(
      .default = col_number(),
      symboling = col_integer(),
      normalized_losses = col_integer(),
      make = col_character(),
      fuel_type = col_character(),
      aspiration = col_character(),
      num_of_doors = col_character(), # Needs manual fix!
      body_style = col_character(),
      drive_wheels = col_character(),
      engine_location = col_character(),
      engine_type = col_character(),
      num_of_cylinders = col_character(), # Also needs manual fix!
      engine_size = col_integer(),
      fuel_system = col_character()
    )
  )
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
## The string-coded numbers are obnoxious; let's map those to integers.
## df_raw %>% pull(num_of_doors) %>% unique()
## df_raw %>% pull(num_of_cylinders) %>% unique()
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
## Manual fixes for some of the variables.
df_autos <-
  df_raw %>%
  mutate(
    num_of_doors = parse_numchar(num_of_doors),
    num_of_cylinders = parse_numchar(num_of_cylinders)
  ) %>%
  rowid_to_column(var = "id")
```

## Define figures for different tasks
See code in `master.Rmd`....

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Common scale
fig_common <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  group_by(drive_wheels, body_style) %>%
  summarize(n = n()) %>%
  ggplot(
    aes(
      x     = body_style,
      y     = n,
      group = drive_wheels,
      color = as.factor(drive_wheels)
    )
  ) +
  geom_line() +
  geom_point(size = 2) +
  scale_color_discrete(name = "Drive") +
  coord_flip() +
  theme_common() +
  ## theme(legend.position = "bottom") +
  labs(
    x = "Style",
    y = "Count",
    title = "COM"
  )
mysave("../images/01_common_scale.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Nonaligned scale
fig_nonaligned <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  ggplot(aes(x = body_style)) +
  geom_bar() +
  coord_flip() +
  ## scale_y_continuous(breaks = c(0, 10, 50)) +
  facet_grid(.~drive_wheels) +
  theme_common() +
  theme(
    aspect.ratio = 4 / 3,
    strip.text.x = element_text(size = 12)
  ) +
  labs(
    x = "Style",
    y = "Count",
    title = "NON"
  )
mysave("../images/02_nonaligned_scale.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Length
fig_length <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  ggplot(aes(x = body_style, fill = as.factor(drive_wheels))) +
  geom_bar() +
  coord_flip() +
  scale_fill_discrete(name = "Drive") +
  theme_common() +
  ## theme(legend.position = "bottom") +
  labs(
    x = "Style",
    y = "Count",
    title = "LEN"
  )
mysave("../images/03_length.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Angle
fig_angle <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  group_by(drive_wheels, body_style) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = "", y = n, fill = drive_wheels)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  scale_fill_discrete(name = "Drive") +
  facet_wrap(.~body_style) +
  theme_common() +
  theme(legend.position = "bottom") +
  labs(
    x = NULL,
    y = "Count",
    title = "ANG"
  )
mysave("../images/04_angle.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Area
fig_area <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  group_by(drive_wheels, body_style) %>%
  summarize(n = n()) %>%
  ggplot(aes(body_style, drive_wheels, size = n)) +
  geom_point() +
  scale_size_continuous(range = c(1, 15)) +
  theme_common() +
  labs(
    x = "Style",
    y = "Drive",
    title = "PIR2"
  )
mysave("../images/05_area.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Color saturation
fig_sat <-
  df_autos %>%
  filter(
    drive_wheels %in% sel_wheels,
    body_style %in% sel_style
  ) %>%
  group_by(drive_wheels, body_style) %>%
  summarize(n = n()) %>%
  ggplot(aes(body_style, drive_wheels, color = n)) +
  geom_point(size = 10) +
  theme_common() +
  labs(
    x = "Style",
    y = "Drive",
    title = "SAT"
  )
mysave("../images/06_sat.png")
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### Package the figures
set.seed(102)
df_figs <-
  tribble(
                 ~task, ~fig,
        "Common scale", fig_common,
    "Nonaligned scale", fig_nonaligned,
              "Length", fig_length,
               "Angle", fig_angle,
                "Area", fig_area,
           "Color sat", fig_sat
  ) %>%
  sample_frac(size = 1)
```

### Plot all visual tasks figures

The following six figures all show identical data -- they each depict a dataset
detailing attributes about different models of car. Each observation in the
dataset is a single model of vehicle; the attributes depicted below are the
Drive style (Forward Wheel Drive (`fwd`) and Rear Wheel Drive (`rwd`)) and Body
style (`hatchback`, `sedan`, or `wagon`). Each figure depicts the number of cars
that fall into each category; for instance, the number of cars in the
dataset that are both `fwd` and `hatchback`.

**Complete the following tasks:**

1. For each visualization, describe *how information is visually encoded.* Is
   information shown via shape, position, color, size, or something else?
2. Answer the question "Which is larger: The count of `wagons` with `fwd` Drive
   OR the count of `wagons` with `rwd` Drive? By how much is one larger than the
   other?"
3. Rank the following visualizations *in terms of how well they help you answer
   question (2) above.* Rate the most helpful visualization as 1, and the least
   helpful as 6; do not allow ties between visualizations.

```{r }
df_figs %>% pull(fig)
```
