---
title: "The Visual Hierarchy"
author: "Zach"
date: 2019-02-16
bibliography: bibtex_database.bib
output: github_document
---

In this lesson, we will illustrate different methods of visualizing information.
The purpose of this exercise is to determine which methods are the most
effective -- in terms of 'ease of interpretation.' You will rank a number of
different visualizations -- of the same data -- in terms of how easily you can
draw comparisons. This will lead to prescriptive suggestions about constructing
visualizations, which will be useful when designing graphics.

<!-- solution-begin -->
The visual hierarchy was proposed by [@cleveland1985graphical] in 1985. The
concept is that certain *perceptual tasks* -- small components of interpreting
graphics -- require more or less effort on the part of the viewer. Easier tasks
are more result in a more readable visualization of data; since our goal should
be to both learn and communicate information as clearly as possible, easier
tasks are better tasks.

Their hierarchy, in order of descending ease, is:

1. Position along a common scale
2. Position on identical but nonaligned scales
3. Length
4. Angle; Slope^[With slope not too close to 0, $\pi/2$, or $\pi$.]
5. Area
6. Volume; Density; Color saturation
7. Color hue
<!-- solution-end -->

```{r setup, echo=FALSE, include=FALSE}
library(tidyverse)

sel_colors <- c("E", "F")
sel_clarities <- c("SI1", "SI2", "VS1", "VS2")

asp_ratio  <- 4 / 4
img_width  <- 5
img_height <- img_width * asp_ratio

# Common theme settings
theme_common <- function() {
  theme_minimal() %+replace%
  theme(
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    axis.title.x = element_text(margin = margin(4, 4, 4, 4), size = 16),
    axis.title.y = element_text(margin = margin(4, 4, 4, 4), size = 16, angle = 90),

    aspect.ratio = asp_ratio,

    ## plot.margin   = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),
    ## panel.spacing = unit(c(t = +0, b = +0, r = +0, l = +0), "cm"),

    legend.text = element_text(size = 10),

    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey90")
  )
}

# Custom save
mysave <- function(filename) {
  ggsave(
    filename = filename,
    height   = img_height,
    width    = img_width,
    units    = "in"
    ## device   = cairo_ps
  )
}
```

<!-- solution-begin -->
## Define figures for different tasks

### Common scale
```{r }
fig_common <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  group_by(clarity, color) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = clarity, y = n, group = color, color = as.factor(color))) +
  geom_line() +
  geom_point(size = 2) +
  scale_color_discrete(name = "Color") +
  coord_flip() +
  theme_common() +
  labs(
    x = "Clarity",
    y = "Count",
    title = "COM"
  )
mysave("../images/01_common_scale.png")
```

### Nonaligned scale
```{r }
fig_nonaligned <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  ggplot(aes(x = clarity)) +
  geom_bar() +
  coord_flip() +
  scale_y_continuous(breaks = c(0, 1e3, 2e3, 3e3)) +
  facet_grid(.~color) +
  theme_common() +
  theme(aspect.ratio = 4 / 3) +
  labs(
    x = "Clarity",
    y = "Count",
    title = "NON"
  )
mysave("../images/02_nonaligned_scale.png")
```

### Length
```{r }
fig_length <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  ggplot(aes(x = clarity, fill = as.factor(color))) +
  geom_bar() +
  coord_flip() +
  scale_fill_discrete(name = "Color") +
  theme_common() +
  labs(
    x = "Clarity",
    y = "Count",
    title = "LEN"
  )
mysave("../images/03_length.png")
```

### Angle
```{r }
fig_angle <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  group_by(clarity, color) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = "", y = n, fill = color)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  scale_fill_discrete(name = "Color") +
  facet_wrap(.~clarity) +
  theme_common() +
  labs(
    x = NULL,
    y = "Count",
    title = "ANG"
  )
mysave("../images/04_angle.png")
```

### Area
```{r }
fig_area <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  group_by(clarity, color) %>%
  summarize(n = n()) %>%
  ggplot(aes(clarity, color, size = n)) +
  geom_point() +
  scale_size_continuous(range = c(1, 15)) +
  theme_common() +
  labs(
    x = "Clarity",
    y = "Color",
    title = "PIR2"
  )
mysave("../images/05_area.png")
```

### Color hue
```{r }
fig_hue <-
  diamonds %>%
  filter(
    color %in% sel_colors,
    clarity %in% sel_clarities
  ) %>%
  group_by(clarity, color) %>%
  summarize(n = n()) %>%
  ggplot(aes(clarity, color, color = n)) +
  geom_point(size = 10) +
  theme_common() +
  labs(
    x = "Clarity",
    y = "Color",
    title = "HUE"
  )
mysave("../images/06_hue.png")
```

### Package the figures
```{r }
set.seed(102)
df_figs <-
  tribble(
                 ~task, ~fig,
        "Common scale", fig_common,
    "Nonaligned scale", fig_nonaligned,
              "Length", fig_length,
               "Angle", fig_angle,
                "Area", fig_area,
           "Color hue", fig_hue
  ) %>%
  sample_frac(size = 1)
```
<!-- solution-end -->

### Plot all visual tasks figures

```{r }
df_figs %>% pull(fig)
```
